---
title: "Data Clean Codes"
author: "Weiwei Song"
output: html_notebook
---
This is an R Markdown Files which specifies how the data of 14 different Indicators for resillence ranking table generated from original multiple csv files. 
The whole model building process are done by WS and YT, including the data searching, data cleaning, the methodology analysis, reactive table building. 
-YT  took charge of 
   search and clean data for several indicators
   edit the methodology_analysis.txt
   optimizes the appearance of reactive table. 
-WS is responsible to 
  collect and clean data for several indicators
  editing the R markdown file data_clean_codes.rmd
  write the codes for the integral part of reactive table(scaled the background color) 
   
  
```{r message=FALSE, warning=TRUE, import,echo=FALSE}
setwd("../data/raw_model_data")
library(tidyverse)
mobility<-read_csv("US_Region_Mobility.csv")
bed<-read_csv("hospital_beds_USA_v1.csv")
chron<-read_csv("chronic_respiratory_bystate.csv")
low<-read_csv("lower_respiratory_infections_mortality_rate_bystate.csv")
lock<-read_csv("lockdown_bystate.csv")
fat<-read_csv("obesity_bystate.csv")
HDI<-read_csv("HDI_USA_2018.csv")
healthcare<-read_csv("healthcare_coverage_rank_2018.csv")
GDP_prediction<-read_csv("US_GDP_prediction_revised.csv")
Dec_data<-read_csv("12-31-2020.csv")
Jan_data<-read_csv("01-31-2021.csv")
vacper100 <- read_csv("us-state-covid-vaccines-per-100.20210212214810444.csv")
vacused <- read_csv("us-share-covid-19-vaccine-doses-used.csv")
vacshare<-read_csv("us-covid-19-share-people-vaccinated.20210212215912385.csv")
total_vac<-read_csv("us-covid-19-total-people-vaccinated.csv")
state_name<-read_csv("state.csv")
```

```{r,Mobility}
#Mobility: take average equally weighted 
mobility_bystate<-
  mobility%>%mutate(sum_mobility=retail_and_recreation_percent_change_from_baseline+
                  grocery_and_pharmacy_percent_change_from_baseline+
                  parks_percent_change_from_baseline+
                  transit_stations_percent_change_from_baseline+
                  workplaces_percent_change_from_baseline+
                  residential_percent_change_from_baseline
                  ) %>%select(sub_region_1,sum_mobility)%>%
  drop_na()%>%
  group_by(sub_region_1)%>%
  summarise(avg=mean(sum_mobility))%>%
  rename(State=sub_region_1,
         Mobility=avg)%>%
 select(State,Mobility)%>%
  drop_na()
```

```{r,HospitalBeds}

HospitalBeds<-bed%>%
  group_by(state)%>%
  summarise(avg=mean(beds))%>%
  ungroup()%>%
  rename(HospitalBeds=avg)%>%
  left_join(.,state_name)%>%
  drop_na()%>%
  select(State,HospitalBeds)

```

```{r,Vaccines}

vac100<-vacper100%>%filter(Date=='2021/2/11')
vacused2<-vacused%>%filter(Date=='2021/2/11')
vacshare2<-vacshare%>%filter(Date=='2021/2/11')
tvac<-total_vac%>%filter(Date=='2021/2/11')
vaccine<-vac100%>%left_join(vacused2,by="Entity")
vaccine<-vaccine%>%left_join(vacshare2,by="Entity")
vaccine<-vaccine%>%left_join(tvac,by="Entity")
vaccine<-vaccine%>%select(Entity, people_vaccinated_per_hundred, total_vaccinations_per_hundred,
                 share_doses_used)
vaccine<-rename(vaccine, c(State=Entity))
```


```{r warning=FALSE}
HDI<-HDI%>%
  rename(State=Region,HDI=`Sub-national HDI`)%>%
  filter(State!="Total")%>%
  select(State, HDI)

Rank_health<-healthcare%>%
  transmute(State=STATE,Rank_health=RANK)%>%
drop_na()

lock<-lock%>%
  rename(State=RegionName)%>%
  drop_na()

GDP_prediction<-GDP_prediction%>%
  transmute(State,GSP=`Gross State Product`)%>%
  drop_na()
  

chron<-chron%>%
  rename(State=Location,chronic_mortality_rate=`Chronic respiratory diseases mortality rate`)%>%
  drop_na()


low<-low%>%
  rename(State=Location,chronic_infection_rate=`Lower respiratory infections mortality rate`)%>%
  drop_na()

```

```{r,covid_columns}
Dec_data<-Dec_data%>%
  transmute(State=Province_State,Confirmed_Dec=Confirmed,Deaths_Dec=Deaths)%>%
  group_by(State)%>%
  summarise_all(sum)

Jan_data<-Jan_data%>%
  transmute(State=Province_State,Confirmed_Jan=Confirmed,Total_Deaths=Deaths,
            Positive_Test_Rate=Confirmed_Jan/Total_Test_Results)%>%
  group_by(State)%>%
  summarise_all(sum)
Covid_columns<-Jan_data%>%
left_join(.,Dec_data,by="State")%>%
  transmute(State,One_Month_Cases=Confirmed_Jan-Confirmed_Dec,
            One_month_fatality_rate=(Total_Deaths-Deaths_Dec)/Total_Deaths,
            Total_Deaths,Positive_Test_Rate)%>%
  drop_na()

```

```{r, all the indictors}
data_indictors<-Covid_columns%>%
  left_join(.,low,by="State")%>%
  left_join(.,chron,by="State")%>%
  left_join(.,GDP_prediction)%>%
  left_join(.,lock,by="State")%>%
  left_join(.,mobility_bystate,by="State")%>%
  left_join(.,Rank_health,by="State")%>%
  drop_na()%>%
  left_join(.,vaccine,by="State")%>%
  left_join(.,HospitalBeds,by="State")%>%
  left_join(.,HDI,by="State")%>%
  left_join(fat)%>%
  rename(assess_to_vaccines_per_hundred=total_vaccinations_per_hundred,
         doses_per_hundred=people_vaccinated_per_hundred,lockdown_severity=lockdown)
summary(data_indictors)
write_csv(data_indictors,"../data/cleaned_model_data/data_indictors.csv")

```

```{r, max_min method}
model_data<-data_indictors
#max-min method function 
I<- function(A){
  min_value<-min(A)
  max_value<-max(A)
  score<-matrix(NA,nrow=length(A))
  for (i in 1:length(A)){
    V<-A[i]
    score[i]<-(V-min_value)/(max_value-min_value)}
  return(score)
}
#use sapply() to all columns 
scale_indicators<-data.frame(sapply(model_data[,-1],I))
scale_indicators$State<-model_data$State
scale_indicators<-scale_indicators%>%
  relocate(State)%>%
  mutate(One_Month_Cases=1-One_Month_Cases,
         One_month_fatality_rate=1-One_month_fatality_rate,
         Total_Deaths=1-Total_Deaths,
         Rank_health=1-Rank_health,
         Positive_Test_Rate=1-Positive_Test_Rate,
         lockdown_severity=1-lockdown_severity,
         chronic_infection_rate
=1-chronic_infection_rate,
         chronic_mortality_rate=1-chronic_mortality_rate, obesity_prevalence=1-obesity_prevalence

)

scale_indicators$Score <- 100*rowMeans(x = scale_indicators[,-1])
scale_indicators$Rank <- 51-rank(scale_indicators$Score)
final_model<-scale_indicators%>%
  select(State,Score,Rank)%>%
  left_join(model_data,.,by="State")
write_csv(scale_indicators,"../data/cleaned_model_data/scale_indictors_of_model.csv")
write_csv(final_model,"../data/cleaned_model_data/final_model.csv")

```

